---
title: "Tidy Data"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE, 
                      comment=NA, 
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align='center')


library(ggplot2)
library(ggthemes)
library(dplyr)
library(tibble)
library(purrr) 
library(tidyverse) 
library(janitor)
library(lubridate)
library(tidyr)
```


## Your Turn 1

The following code creates the dataset used in this exercise. 

```{r}
BP_wide <- tibble(id = letters[1:4],
                     sex = c("F", "M", "M", "F"),
                     SBP_v1 = c(130, 120, 130, 119),
                     SBP_v2 = c(110, 116, 136, 106),
                     SBP_v3 = c(112, 122, 138, 118))
BP_wide
```

### a. What do you think the data in the table are measures of? How can we tell the data are wide?

*Answer:*  


### b. What would a long data look like? How many columns will it have?

*Answer:* 


----------------------------------------

Your Turn 2

```{r}
DBP_wide <- tibble(id = letters[1:4],
 sex = c("F", "M", "M", "F"),
 v1.DBP = c(88, 84, 102, 70),
 v2.DBP = c(78, 78, 96, 76),
 v3.DBP = c(94, 82, 94, 74),
 age=c(23, 56, 41, 38)
 )
DBP_wide

```


```{r}
BP_long <- BP_wide %>% 
  gather(key = "visit", value = "SBP", SBP_v1:SBP_v3)
BP_long2 <- BP_long %>% 
  mutate(visit = str_replace(visit,"SBP_v",""))
BP_long2
```


### a. Make `DBP_wide` into a long dataframe based on the repeated DBP columns and save it as `DBP_long`

```{r, eval=FALSE}
DBP_long <- DBP_wide %>%
  gather(key = _____", value = _____, _______)
DBP_long
```


```{r}
DBP_long <- DBP_wide %>%
  pivot_longer(names_to = _____,
               values_to = ______,
               cols = ______)
DBP_long
```


### b. Clean up the visit column of `DBP_long` so that the values are 1, 2, 3, and save it as `DBP_long`

```{r}
DBP_long <- DBP_long %>%
  mutate(visit = str_replace(_________)) %>%
  mutate(visit = str_replace(_________))  
DBP_long
```


### c. Make `DBP_long` wide with column names visit.1, visit.2, visit.3 for the DBP values, and save it as `DBP_wide2`

```{r, eval=FALSE}
DBP_wide2 <- DBP_long %>% 
  spread(key = _____, value = _____, sep=".") # specify separating character
DBP_wide2
```


```{r}
DBP_wide2 <- DBP_long %>% 
  pivot_wider(names_from = ____,
              values_from = ____)
DBP_wide2
```


### d. Join `DBP_long` with `BP_long2` so that we have one data frame with columns id, sex, visit,
SBP, DBP, and age. Save this as `BP_both_long`.


```{r}
BP_both_long <- left_join(______, ______, by = c(________________))
BP_both_long
```

---------------------------------------------


# Your Turn 3

```{r, echo=FALSE}
messy_data <- tibble(NAME = c("J N","A C","D E"), 
                     `months follow up` = c("", 10, 11), 
                     `Date of visit` = c("July 31, 2003", "Nov 12, 2005", "Aug 3, 2007"))

```



### a. Clean column names with clean_names(). Store the clean data under `clean_data`.

```{r}
clean_data <- _________ %>% 
  clean_names()
clean_data
```


### b. Replace missing ("") data in months_follow_up with NA.

```{r}
clean_data <- messy_data %>% 
  clean_names() %>%
  mutate(
    months_follow_up = replace_na(________________))
clean_data
```



### c. Convert months_follow_up to a numeric variable.


```{r}
clean_data <- messy_data %>% 
  clean_names() %>%
  mutate(
    months_follow_up = replace_na(_____________),
    months_follow_up = as.numeric(_____________))
clean_data
```


### d. Convert date_of_visit to a date.

```{r}
clean_data <- messy_data %>% 
  clean_names() %>%
  mutate(
    months_follow_up = replace_na(_____________),
    months_follow_up = as.numeric(_____________),
    date_of_visit = mdy(____________))
clean_data

```


### e. Create a column called date_last_visit that is the date of visit plus months of follow up.


```{r}
clean_data <- messy_data %>% 
  clean_names() %>%
  mutate(
    months_follow_up = replace_na(_____________),
    months_follow_up = as.numeric(______________),
    date_of_visit = mdy(____________),
    date_last_visit = __________ + ______________)
clean_data
```



### f. Remove rows (cases) with missing data in months_follow_up.

```{r}
clean_data <- clean_data %>%
  drop_na(______________)
clean_data
```



### g. Remove the spaces in name.

```{r}
clean_data <- clean_data %>%
  drop_na(_____________) %>%
  mutate(name = __________________)
clean_data
```


